---
title: "Calculating Customer Price Sensitivity Using Discrete Choice Model and Market-level Data "
author: "Yurong Fan""
date: "03/04/2017"
---

# Data Processing
## Merge catsales with zone by store
```{r}
library(dplyr)
library(ggplot2)
library(AER)
library(stargazer)

#load data and merge
catsales<-read.csv("catsales_RefrigOJ_2017Semester.csv")
zone <- read.csv("zone.csv")

catsales <- catsales%>%
  left_join(zone, by = "store")%>%
  mutate(zone = as.factor(zone))
```

#Comparing Statistics for Given and Not Given Zones
```{r}
givenzones <- catsales[!is.na(catsales$zone),c("reprice","resales","move","prom","whprice")]
nazones <- catsales[is.na(catsales$zone),c("reprice","resales","move","prom","whprice")]
```

```{r}
ggplot() + geom_density(aes(x =reprice, fill= "blue"), alpha= 0.5, data = givenzones) + 
  geom_density(aes(x = reprice, fill = "green"), alpha=0.5, data = nazones) + 
  scale_fill_discrete(name = "Zones",labels=c("Not Given", "Given")) + 
  xlab("Price") + ylab("Distribution")

ggplot() + geom_density(aes(x =resales, fill= "blue"), alpha= 0.5, data = givenzones) + 
  geom_density(aes(x = resales, fill = "green"), alpha=0.5, data = nazones) + 
  xlim(0,300) + scale_fill_discrete(name= "Zones", labels=c("Not Given", "Given")) + 
  xlab("Sales") + ylab("Distribution")
```


#Removes rows where store has NA in zone column
```{r}
catsales = filter(catsales, !is.na(zone))
```

#Exploratory Data Analysis
## Basic questions
```{r}
#how many weeks? 
length(unique(catsales$week))

#how many stores?
length(unique(catsales$store))

#how many brands and products for each brand
upc_per_brand = catsales%>%
  group_by(unibranddff)%>%
  summarise(upc_num = n_distinct(upc))%>%
  ungroup()

upc_per_brand
```



## Market Share
```{r}
#market share  
whole_catsale <- catsales%>%
  group_by(week)%>%
  summarise(whole_cat_sales = sum(resales))

byzone_catsale <- catsales%>%
  group_by(week,zone)%>%
  summarise(zone_cat_sales = sum(resales))

catsales <- catsales%>%
  left_join(whole_catsale, by = "week")%>%
  left_join(byzone_catsale, by = c("week","zone"))%>%
  mutate(byzone_market_share = resales/zone_cat_sales * 100)%>%
  mutate(whole_market_share = resales/whole_cat_sales * 100)

# market share by brand in the whole market
brand_mkt_shr = catsales%>%
  group_by(unibranddff,week)%>%
  summarise(market_share_bybrand = sum(whole_market_share))%>%
  group_by(unibranddff)%>%
  summarise(avg_market_share_bybrand = round(mean(market_share_bybrand),2))%>%
  arrange(desc(avg_market_share_bybrand))

# by brand market share in the whole market over time
brand_share_overtime <- catsales%>%
  group_by(unibranddff,week)%>%
  summarise(brand_market_share = sum(whole_market_share))


# market share by brand by zone
catsales%>%
  group_by(unibranddff,week,zone)%>%
  summarise(market_share_bybrand = sum(byzone_market_share))%>%
  group_by(unibranddff,zone)%>%
  summarise(avg_market_share_bybrand = mean(market_share_bybrand))%>%
  arrange(desc(zone, avg_market_share_bybrand))

```


#Creates weighted price, price per ounce and price per move(number of actual items sold) columns
```{r}
catsales$weight_reprice = catsales$reprice*catsales$resales
catsales$reprice_move = catsales$reprice*catsales$move
```


#Aggregate Data
```{r}

#further aggregate by brand using simple average for price
aggcatsales <- catsales%>%
  group_by(zone,week,unibranddff)%>%
  summarise(reprice = mean(reprice),
            resales = sum(resales),
            prom = mean(prom),
            whprice = mean(whprice),
            move = sum(move),
            market_share = sum(byzone_market_share),
            weight_reprice = sum(weight_reprice)/sum(resales),
            reprice_move = sum(reprice_move)/sum(move))%>%
  ungroup()


#transform baseline 
other_share <- aggcatsales%>%
  filter(unibranddff == "Other")%>%
  select(zone, week, market_share)

names(other_share)[3] = "other_share"

aggcatsales <- aggcatsales%>%
  left_join(other_share, by = c('zone', 'week'))
```


#Build Aggregate-level Demand Model To Calculate Customer Price Sensitivity
#Change price messurements for model robustness check 

## Using OLS with price and promotion as product characteristics
```{r}
#generate dependent variable in dcm 
aggcatsales$Y = log(aggcatsales$market_share/aggcatsales$other_share, base = exp(1))

#Using OLS with price and promotion as product characteristics
dcm1 <- lm(data = aggcatsales, Y ~ reprice + prom)
```

## Using OLS with price and promotion as product characteristics and brand dummies.
```{r}
dcm2 <- lm(data = aggcatsales, Y ~ reprice + prom + factor(unibranddff))
```

## Using OLS with price and promotion as product characteristics and brand and zone dummies.
```{r}
dcm3 <- lm(data = aggcatsales, Y ~ reprice + prom 
           + factor(unibranddff) + factor(zone))
```

##Estimate the models of 1, 2 and 3 using wholesale costs as an instrument for price.
```{r}
dcm4 <- ivreg(Y ~ reprice + prom | whprice + prom, data = aggcatsales)
dcm5 <- ivreg(Y ~ reprice + prom + factor(unibranddff) | whprice + prom 
              + factor(unibranddff), data = aggcatsales)
dcm6 <- ivreg(Y ~ reprice + prom + factor(unibranddff) 
              + factor(zone) | whprice + prom + factor(unibranddff) 
              + factor(zone), data = aggcatsales)
```

##Estimate the models of 1, 2 and 3 using Hausman instruments (average price in other zones)
```{r}
# generate average price in other zones
IV_other_zone_price <- aggcatsales%>%
  group_by(week, unibranddff)%>%
  summarise(tot_price = sum(reprice),
            num_zone = n_distinct(zone))%>%
  left_join(aggcatsales, by = c("week", "unibranddff"))%>%
  mutate(other_zone_price = (tot_price - reprice)/(num_zone - 1))%>%
  select(week, unibranddff, zone, other_zone_price )

aggcatsales <- aggcatsales%>%
  left_join(IV_other_zone_price, by = c("week", "unibranddff","zone"))


dcm7 <- ivreg(Y ~ reprice + prom | other_zone_price + prom, data = aggcatsales)
dcm8 <- ivreg(Y ~ reprice + prom + factor(unibranddff) | other_zone_price 
              + prom + factor(unibranddff), data = aggcatsales)
dcm9 <-ivreg(Y ~ reprice + prom + factor(unibranddff) + factor(zone) | other_zone_price 
             + prom + factor(unibranddff) + factor(zone), data = aggcatsales)
```

## Summarize Model Result In One Table
```{r}
stargazer(dcm1,dcm2,dcm3,dcm4,dcm5,dcm6,dcm7,dcm8,dcm9, 
          type = "text", 
          omit = c("zone","Constant"), 
          covariate.labels=c("reprice", "prom", "FLDAGOLD", "FLORIDA", "HH", "MMAID", "Other", "TROP"),
          out = "model_123.txt", float.env = "sidewaystable")

```

