---
title: "Evaluating Property Values's Effect on Water Quality Using Panel Data"
output: html_document
author: Yurong Fan
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Preperation
```{r, echo = FALSE}
setwd('C:/Users/Fanfan/Desktop/Water/data')
lakeNeighbor = read.csv("allLakeNeighborhood_panel_dt_v6_1000_TP.csv")
library(dplyr)
library(data.table)
library(ggplot2)
library(plm)

#merge with external data 
lake_metadata = read.csv('lake_metadata.csv',stringsAsFactors=FALSE)
lakeNeighbor = left_join( lakeNeighbor, lake_metadata, by = c("pw_MCES_Map_Code1" = "DNR_ID_SITE_NUMBER"))
weather = read.csv('weather.csv',stringsAsFactors=FALSE)
lakeNeighbor = left_join( lakeNeighbor, weather, by = c("COUNTY", "p_year" = "YEAR" ))

#select useful columns
lakeNeighbor = lakeNeighbor%>%
  select(LAKE_NAME, p_year, MAJOR_WATERSHED, COUNTY, EMVTotal, EMV_Total_res, EMV_Total_com, EMV_Total_ind, EMV_Total_agr, EMV_Total_agr, EMV_Total_sch, EMV_Total_hos, EMV_Total_fed, EMV_Total_greenN, EMV_Total_greenY, avgTP, avgDepth, avgGrade, avgRec, PRCP,nearwinter_snow, pw_MCES_Map_Code1)

lakeNeighbor$entity = as.factor(lakeNeighbor$pw_MCES_Map_Code1)
lakeNeighbor$year = as.factor(lakeNeighbor$p_year)

#create lagged term of the target variable 
lakeNeighbor = lakeNeighbor %>%
  group_by(entity) %>%
  mutate(avgTP_lag = lag(avgTP, 1))%>%
  mutate(avgDepth_lag = lag(avgDepth, 1))%>%
  mutate(avgGrade_lag = lag(avgGrade, 1))%>%
  mutate(avgRec_lag = lag(avgRec, 1))
attach(lakeNeighbor)
```

#Data Summary
```{r}
#non-zero rows by attributs
colSums(lakeNeighbor != 0)
```
residential, commercial, argriculture has more non-zero rows
industrial and school has some non-zero rows, others user type has too few oberservations 

```{r}
cor(cbind(EMVTotal_w, EMV_Total_agr, EMV_Total_res, EMV_Total_com, EMV_Total_ind, EMV_Total_greenN, EMV_Total_greenY), use = "pairwise.complete.obs")
```
EMV_Total_greenN and EMVTotal, 
EMV_Total_greenY and EMV_Total_agr,
EMVLand, EMVBldg and EMVTotal,
EMVLand_sqt, EMVTotal_sqt,
EMV_Total_greenN and EMVBldg,
counts and EMVTotal, EMVLand, EMVBuilding,
has high correlation.


# Cleaning outliers
```{r}
#plot distribution
ggplot(lakeNeighbor, aes(EMVTotal_w)) + geom_density()
ggplot(lakeNeighbor, aes(EMV_Total_agr)) + geom_density()
ggplot(lakeNeighbor, aes(avgTP)) + geom_density()
ggplot(lakeNeighbor, aes(PRCP)) + geom_density()
ggplot(lakeNeighbor, aes(nearwinter_snow)) + geom_density()

#clear outliers
outlierReplace = function(dataframe, cols, rows, newValue = NA) {
  if (any(rows)) {
    set(dataframe, rows, cols, newValue)
  }
}
lakeNeighbor = data.table(lakeNeighbor)

ggplot(lakeNeighbor, aes(avgTP)) + geom_density()
outlierReplace(lakeNeighbor, "avgTP", which(avgTP > 1), NA)
ggplot(lakeNeighbor, aes(EMVTotal_w)) + geom_density()
outlierReplace(lakeNeighbor, "EMVTotal_w", which(EMVTotal_w > 1.0e+07), NA)
ggplot(lakeNeighbor, aes(EMV_Total_agr)) + geom_density() 
outlierReplace(lakeNeighbor, "EMV_Total_agr", which(EMV_Total_agr >70000), NA)
ggplot(lakeNeighbor, aes(EMV_Total_res)) + geom_density() 
outlierReplace(lakeNeighbor, "EMV_Total_res", which(EMV_Total_res >7000000), NA)
ggplot(lakeNeighbor, aes(EMV_Total_com)) + geom_density() 
outlierReplace(lakeNeighbor, "EMV_Total_com", which(EMV_Total_com >500000), NA)
ggplot(lakeNeighbor, aes(EMV_Total_ind)) + geom_density() 
outlierReplace(lakeNeighbor, "EMV_Total_ind", which(EMV_Total_ind >3332700), NA)
ggplot(lakeNeighbor, aes(EMV_Total_greenY)) + geom_density() 
outlierReplace(lakeNeighbor, "EMV_Total_greenY", which(EMV_Total_greenY >70000), NA)
```


#filter out lakes with more than 6 years data 
```{r, echo = FALSE}
efilter = lakeNeighbor%>%
  group_by(entity)%>%
  summarise(n=n())%>%
  filter(n>6)
lakeNeighbor = inner_join(lakeNeighbor,efilter, on = "entity")
```

#use EMVTotal as the predictor and test fix effects
```{r}
#test poolability
pool = plm(avgTP ~ avgTP_lag + EMVTotal_w  + PRCP + nearwinter_snow, data = lakeNeighbor, index = c("entity", "year"), model = "pooling")
fix_e = plm(avgTP ~ avgTP_lag + EMVTotal_w + PRCP + nearwinter_snow, data = lakeNeighbor, index = c("entity", "year"), model = "within", effect = "individual")

pooltest(pool, fix_e)
##pool model is not suitable 


#test individual and time effect
plmtest(pool, effect = "individual")
## individual effect is significant
plmtest(pool, effect = "time")
## time effect is significant 
plmtest(pool, effect = "twoway")
##twoway effect is significant 

#test random effect 
fix = plm(avgTP ~ avgTP_lag + EMVTotal_w  + PRCP + nearwinter_snow, data = lakeNeighbor, index = c("entity", "year"), model = "within")
random = plm(avgTP ~ avgTP_lag + EMVTotal_w  + PRCP + nearwinter_snow, data = lakeNeighbor, index = c("entity", "year"), model = "random")
phtest(fix, random)
##fix effect is better than random effect

#test serial correlation
pwtest(avgTP ~ EMVTotal_w  + PRCP + nearwinter_snow, data = lakeNeighbor)
##there are unobserved individual effect that may introduce bias when using fixed effect
pwtest(avgTP ~ avgTP_lag + EMVTotal_w  + PRCP + nearwinter_snow, data = lakeNeighbor)
##add lag term of the dependant variable as instrument alleviated unobserved individual effect
```

#use two way fixed effect to test the effect of EMVTotal on avgTP
```{r}
fix_te = plm(avgTP ~ avgTP_lag + EMVTotal_w  + PRCP + nearwinter_snow, data = lakeNeighbor, index = c("entity", "year"), model = "within", effect = "twoways")
summary(fix_te)

fix_te = plm(avgDepth ~ avgDepth_lag + EMVTotal_w  + PRCP + nearwinter_snow, data = lakeNeighbor, index = c("entity", "year"), model = "within", effect = "twoways")
summary(fix_te)

fix_te = plm(avgGrade ~ avgGrade_lag + EMVTotal_w  + PRCP + nearwinter_snow, data = lakeNeighbor, index = c("entity", "year"), model = "within", effect = "twoways")
summary(fix_te)

fix_te = plm(avgRec ~ avgRec_lag + EMVTotal_w  + PRCP + nearwinter_snow, data = lakeNeighbor, index = c("entity", "year"), model = "within", effect = "twoways")
summary(fix_te)
```

#By use type EMV's effect on avgTP
## all in one
```{r}
fix_te = plm(avgTP ~ avgTP_lag + EMVTotal_w + EMV_Total_agr + EMV_Total_com + EMV_Total_ind + EMV_Total_res + PRCP + nearwinter_snow, data = lakeNeighbor, index = c("entity", "year"), model = "within", effect = "twoways")
summary(fix_te)
```
industrial negative 

##Agriculture
```{r}
Ln_agr = lakeNeighbor%>%filter(EMV_Total_agr > 0)

fix_te = plm(avgTP ~ avgTP_lag + EMV_Total_agr + PRCP + nearwinter_snow, data = Ln_agr, index = c("entity", "year"), model = "within", effect = "twoways")
summary(fix_te)

fix_te = plm(avgDepth ~ avgDepth_lag + EMV_Total_agr + PRCP + nearwinter_snow, data = Ln_agr, index = c("entity", "year"), model = "within", effect = "twoways")
summary(fix_te)
```


##residential
```{r}
Ln_res = lakeNeighbor%>%filter(EMV_Total_res > 0)

fix_te = plm(avgTP ~ avgTP_lag + EMV_Total_res + PRCP + nearwinter_snow, data =  Ln_res, index = c("entity", "year"), model = "within", effect = "twoways")
summary(fix_te)

fix_te = plm(avgDepth ~ avgDepth_lag + EMV_Total_res + PRCP + nearwinter_snow, data =  Ln_res, index = c("entity", "year"), model = "within", effect = "twoways")
summary(fix_te)
```


##commercial
```{r}
Ln_com = lakeNeighbor%>%filter(EMV_Total_com > 0)

fix_te = plm(avgTP ~ avgTP_lag + EMV_Total_com + PRCP + nearwinter_snow, data =  Ln_com, index = c("entity", "year"), model = "within", effect = "twoways")
summary(fix_te)

fix_te = plm(avgDepth ~ avgDepth_lag + EMV_Total_com + PRCP + nearwinter_snow, data =  Ln_com, index = c("entity", "year"), model = "within", effect = "twoways")
summary(fix_te)
```

##industrial
```{r}
Ln_ind = lakeNeighbor%>%filter(EMV_Total_ind > 0)

fix_te = plm(avgTP ~ avgTP_lag + EMV_Total_ind + PRCP + nearwinter_snow, data =  Ln_ind, index = c("entity", "year"), model = "within", effect = "twoways")
summary(fix_te)

fix_te = plm(avgDepth ~ avgDepth_lag + EMV_Total_ind + PRCP + nearwinter_snow, data =  Ln_ind, index = c("entity", "year"), model = "within", effect = "twoways")
summary(fix_te)
```

##greenY
```{r}
fix_te = plm(avgTP ~ avgTP_lag + EMV_Total_greenY + PRCP + nearwinter_snow, data =  lakeNeighbor, index = c("entity", "year"), model = "within", effect = "twoways")
summary(fix_te)

fix_te = plm(avgDepth ~ avgDepth_lag + EMV_Total_greenY + PRCP + nearwinter_snow, data =  lakeNeighbor, index = c("entity", "year"), model = "within", effect = "twoways")
summary(fix_te)
```

#By county
##Create by county subset
```{r}
lakeNeighbor%>%group_by(COUNTY)%>%summarise(n=n())

Carver = lakeNeighbor%>%filter(COUNTY == "Carver")
Dakota = lakeNeighbor%>%filter(COUNTY == "Dakota")
Hennepin = lakeNeighbor%>%filter(COUNTY == "Hennepin")
Washington = lakeNeighbor%>%filter(COUNTY == "Washington")
Scott = lakeNeighbor%>%filter(COUNTY == "Scott")
```

##Carver
```{r}
ggplot() + geom_point(data = Carver, aes(x = year, y = avgTP,na.rm = TRUE))
ggplot() + geom_point(data = Carver, aes(x = year, y = EMVTotal_w))
ggplot() + geom_point(data = Carver, aes(x = year, y = EMV_Total_ind))
ggplot() + geom_point(data = Carver, aes(x = year, y = EMV_Total_com))
ggplot() + geom_point(data = Carver, aes(x = year, y = EMV_Total_res))

fix_te = plm(avgTP ~ avgTP_lag + EMVTotal_w + EMV_Total_agr + EMV_Total_com + EMV_Total_ind + EMV_Total_res + PRCP + nearwinter_snow, data = Carver, index = c("entity", "year"), model = "within", effect = "twoways")
summary(fix_te)
```



#By Watershed
##Create by county subset
```{r}
lakeNeighbor%>%group_by(MAJOR_WATERSHED)%>%summarise(n=n())

LMR = lakeNeighbor%>%filter(MAJOR_WATERSHED == "Lower Minnesota River")
LSCR = lakeNeighbor%>%filter(MAJOR_WATERSHED == "Lower St. Croix River")
MRTC = lakeNeighbor%>%filter(MAJOR_WATERSHED == "Mississippi River - Twin Cities")
```

## Mississippi River - Twin Cities
```{r}
summary(MRTC)
ggplot() + geom_point(data = MRTC, aes(x = year, y = avgTP))
ggplot() + geom_point(data = MRTC, aes(x = year, y = EMVTotal_w))
ggplot() + geom_point(data = MRTC, aes(x = year, y = EMV_Total_ind))
ggplot() + geom_point(data = MRTC, aes(x = year, y = EMV_Total_com))
ggplot() + geom_point(data = MRTC, aes(x = year, y = EMV_Total_res))

fix_te = plm(avgTP ~ avgTP_lag + EMVTotal_w + EMV_Total_agr + EMV_Total_com + EMV_Total_ind + EMV_Total_res + PRCP + nearwinter_snow, data = MRTC, index = c("entity", "year"), model = "within", effect = "twoways")
summary(fix_te)
```

